**Setup**
```{r setup}
# Set working directory
library(here)
here()

# Load packages
library(tidyverse)
library("dataRetrieval")
library(dplyr)
library(lubridate)
library(ggplot2)
```

**Explore using plots**
*Mean Daily Discharge (cfs)*
```{r}
ggplot(FBR_Hotsprings.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03454500, French Broad River at Hot Springs, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)") 

ggplot(FBR_Marshall.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03453500, French Broad River at Marshall, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(FBR_Asheville.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03451500, French Broad River at Asheville, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(BeetreeCreek_Swannanoa.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03450000, Bee Tree Creek near Swannanoa, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(NFSwannanoa_Walkertown.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #0344894205, North Fork Swannanoa River near Walkertown, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(Swannanoa_Biltmore.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03451000, Swannanoa River at Biltmore, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(FBR_Fletcher.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03447687, French Broad River near Fletcher, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(FBR_Blantyre.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03443000, French Broad River at Blantyre, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")

ggplot(FBR_Rosman.Dailydischarge.combined, aes(x = date, y = mean_daily_discharge_cfs)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Daily Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03439000, French Broad River at Rosman, NC",
       x = "Sept 27 - March 27",
       y = "Mean Daily Discharge (cfs)")
```

*Mean Monthly Discharge*
```{r}
month_order <- c(9:12, 1:3)
month_order_labels <- month.abb[month_order]

ggplot(FBR_Hotsprings.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order, 
                      labels = month_order_labels), y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03454500, French Broad River at Hot Springs, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(FBR_Marshall.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels), 
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03453500, French Broad River at Marshall, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(FBR_Asheville.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels), 
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03451500, French Broad River at Asheville, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(BeetreeCreek_Swannanoa.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels), 
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03450000, Bee Tree Creek near Swannanoa, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(NFSwannanoa_Walkertown.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels),
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #0344894205, North Fork Swannanoa River near Walkertown, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(Swannanoa_Biltmore.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels),
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03451000, Swannanoa River at Biltmore, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(FBR_Fletcher.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels), 
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03447687, French Broad River near Fletcher, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(FBR_Blantyre.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels), 
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03443000, French Broad River at Blantyre, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")

ggplot(FBR_Rosman.Monthlydischarge.combined, 
       aes(x = factor(month, levels = month_order,
                      labels = month_order_labels), 
           y = mean_monthly_discharge_cfs)) +
  geom_line(group = 1) +
  geom_smooth(method = "lm") +
  facet_wrap(~time_period, scales = "free_x", ncol = 2) +
  labs(title = "Mean Monthly Discharge (cfs) before and after Hurricane Helene",
       subtitle = "USGS Gage #03439000, French Broad River at Rosman, NC",
       x = "Sept 27 - March 27",
       y = "Mean Monthly Discharge (cfs)")
```

*Monthly Discharge Trends by Site (Before vs. After Hurricane Helene)*
```{r}
# Combine all monthly discharge datasets into one
monthly_combined_all_sites <- bind_rows(
  FBR_Hotsprings.Monthlydischarge.combined,
  FBR_Marshall.Monthlydischarge.combined,
  FBR_Asheville.Monthlydischarge.combined,
  BeetreeCreek_Swannanoa.Monthlydischarge.combined,
  NFSwannanoa_Walkertown.Monthlydischarge.combined,
  Swannanoa_Biltmore.Monthlydischarge.combined,
  FBR_Fletcher.Monthlydischarge.combined,
  FBR_Blantyre.Monthlydischarge.combined,
  FBR_Rosman.Monthlydischarge.combined
)

# Define month order for plotting (September–March)
month_order <- c(9:12, 1:3)
month_order_labels <- month.abb[month_order]

# Create a multi-facet plot to visualize all gauge sites
ggplot(monthly_combined_all_sites, 
       aes(x = factor(month, levels = month_order, labels = month_order_labels), 
           y = mean_monthly_discharge_cfs,
           group = time_period,
           color = time_period)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~ name, scales = "free_y") +
  labs(
    title = "Monthly Mean Discharge Before and After Hurricane Helene",
    subtitle = "All USGS Gauge Sites in the French Broad River Basin",
    x = "Month (Sept–March)",
    y = "Mean Monthly Discharge (cfs)",
    color = "Period"
  ) +
  theme_minimal(base_size = 13) +
  theme(strip.text = element_text(face = "bold"))
```
*Summary Table of Before and After Mean Monthly Discharge by Site*
```{r}
# Combine all monthly datasets
monthly_combined <- bind_rows(
  FBR_Hotsprings.Monthlydischarge.combined,
  FBR_Marshall.Monthlydischarge.combined,
  FBR_Asheville.Monthlydischarge.combined,
  BeetreeCreek_Swannanoa.Monthlydischarge.combined,
  NFSwannanoa_Walkertown.Monthlydischarge.combined,
  Swannanoa_Biltmore.Monthlydischarge.combined,
  FBR_Fletcher.Monthlydischarge.combined,
  FBR_Blantyre.Monthlydischarge.combined,
  FBR_Rosman.Monthlydischarge.combined
)

# Summary table: mean by site and period
discharge_summary <- monthly_combined %>%
  group_by(name, ID_number, time_period) %>%
  summarise(mean_discharge = mean(mean_monthly_discharge_cfs, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = time_period, 
              values_from = mean_discharge) %>%
  mutate(
    difference_cfs = `2024-2025` - `2023-2024`,
    percent_change = (`2024-2025` - `2023-2024`) / `2023-2024` * 100
  )

# View result
print(discharge_summary)
```
The table compares mean monthly discharge before and after Hurricane Helene at nine USGS gauge sites in the French Broad River Basin.

Discharge increased at all sites after the hurricane.
The percent change ranged from +163% at Rosman (USGS 03439000) to +926% at Swannanoa Biltmore (USGS 03451000),
with the largest increases observed at sites with lower pre-hurricane discharge.

These results suggest that Hurricane Helene affected local hydrology, likely increasing runoff and streamflow in the following months. They also highlight how extreme weather events can alter river conditions, especially in areas with varied terrain and land use.
